---
title: "Statistical Analysis of Employee Satisfaction Data"
author: "林渭倫 R13227114"
output: html_document
date: "`r format(Sys.time(), '%Y-%m-%d')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Background

- **Author**: 林渭倫
- **Created At**: 2025-10-20
- **Research Motivation and Context (why are we interested in the findings?)：**
- **Main Findings and Takeaways：**
- **Future Direciton：**

```{r load-packages}
# Load packages here
library(tidyverse)
library(kableExtra)
library(MASS)
```

```{r load-data}
# Load input data here and please finish all the data manipulation here.
# Finish this block by printing the first ten observations of the data.
# Note:
# - You may only read data from /data/processed.
# - Files in /data/processed should already be cleaned and prepared for analysis.
# - Beyond simple filtering of observations or generating a small number of variables,
#   further data manipulation is not allowed. If more extensive changes are needed,
#   update the source data instead.
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))  # set the working directory to this Rmd file
dat <- read.csv("../data/processed/processed_data112.csv")

dat <- dat %>% 
  mutate_at(-c(14, 29, 30), as.factor)  # convert mislabeled factor into actual factor
head(dat, 10) %>% 
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```

```{r}
# Before conducting the analysis, report summary statistics for all variables 
# that will be used in this notebook:
#   - For numeric variables: mean, min, 25th percentile, median, 75th percentile, 
#     max, standard deviation, and number of observations.
#   - For categorical variables: number of observations and the counts/percentages 
#     of the ten most frequent categories.
# Conclude this block with the complete summary statistics.
```

### The actual analysis starts below
Make the graphs, summary statistics, regression model below.
Make sure you have followed the guidelines as specified in 
[專案資料夾結構、檔案命名與文件規範](https://docs.google.com/document/d/1sl6gEFMdmiGsiNjLe17UmZ30xKxq15U0Mb2B-Jvusxg/edit?tab=t.33iie8ybx7s4).

The summary statistics for numeric variables:

```{r summary-stat-numeric}
dat %>% 
  rename_with(~ gsub("_", ".", .x, fixed=T)) %>%   # replace the "_" in the column name. should be fixed after naming of columns
  summarise(across(-where(is.factor), .fns=list(
    mean = mean,
    min = min,
    pr.25th = ~ quantile(., .25),
    median = median,
    pr.75th = ~ quantile(., .75),
    max = max,
    stdev = sd,
    n = ~ length(.)
  ))) %>% 
  pivot_longer(everything(), names_sep = "_", names_to = c("variable", ".value")) %>% 
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```

The histograms of the numeric variables:

```{r plot-numeric}
dat %>% 
  select(-is.factor) %>% 
  pivot_longer(everything()) %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~ name, scales="free") +
  labs(title = "Histograms of Numeric Variables") +
  theme_light()
```

The summary statistics for categorical variables:

```{r summary-stat-catogorical}
cate_summary <- dat %>% 
  # rename_with(~ gsub("_", ".", .x, fixed=T)) %>%   # replace the "_" in the column name. should be fixed after naming of columns
  summarise(across(where(is.factor), .fns=list(
    # n = ~ length(.),
    level = ~ list(table(.))
  ))) %>% 
  pivot_longer(everything(), names_sep = "_", names_to = c("variable", ".value")) %>%   # arrange to #obs and levels
  unnest_wider(level) %>%  # unnest the count of levels 
  mutate(across(num_range("", 0:20), as.numeric)) %>%   # change the data type of count
  pivot_longer(  # sort columns of levels to a single column `count` and its naming `level`
    num_range("", 0:20),
    names_to = "level", values_to = "count"
  ) %>% 
  drop_na() %>%  # remove excess levels
  group_by(variable) %>% 
  mutate(
    percentage = (100 * count / sum(count)) %>%  # compute percentage across different variables
      round(2) %>% 
      paste0("%")
  ) %>% 
  arrange(variable, desc(count)) %>%  # sorting by variable
  filter(row_number() <= 10)  # only focus on the ten most frequent categories

cate_summary %>% 
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>% 
  collapse_rows(1) %>% 
  footnote(general = "All the numbers of observation of categorical variable are 4095")
```

We first defined a function that helps drawing pie chart of categorical variables.

```{r plot-categorical}
piechart <- function(dat, var.range=1:9) {
  dat %>%
    filter(variable %in% unique(.$variable)[var.range]) %>%
    # arrange(variable, count) %>% 
    ggplot(aes(x="", y=count, fill=level)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) +
    theme_void() +
    geom_text(
      # aes(label = paste0(count, " (", percentage, ")")),
      aes(label = level),
      size = 3, position = position_stack(vjust = .5)
    ) +
    facet_wrap(~ variable, scales="free", drop=T) +
    labs(title = "Piecharts of Categorical Variables")
}

n.chart <- 6
for (i in seq(1, 27, n.chart)) {
  print(piechart(cate_summary, i:(i + n.chart - 1)))
}
```

Ordinal logistic regression

```{r olr}
mod <- polr(a12 ~ ., data=dat)

summary(mod)
```

```{r coef}
mod.coef <- coef(summary(mod)) %>% 
  as.data.frame() %>% 
  rename_with(~ gsub(" ", ".", .x, fixed=T)) %>% 
  rownames_to_column("variables") %>% 
  mutate(
    p.value = pnorm(abs(t.value), lower.tail=F)*2
  ) %>% 
  mutate(across(-1, ~ round(., 3)))

mod.confint <- confint.default(mod) %>% 
  as.data.frame() %>% 
  round(3) %>% 
  rename_with(~ gsub(" %", "% CI", .x)) %>% 
  rownames_to_column("variables")

left_join(mod.coef, mod.confint) %>% 
  mutate(
    p.value = cell_spec(
      p.value, "html", 
      color=ifelse(p.value < .05, "red", "black")
    )
  ) %>% 
  kable("html", escape=F) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```



